---
title: "3. OSCA Human Improved Analysis"
author: "Dario Righelli"
date: "`r Sys.Date()`"
output: html_document
output: 
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{RobinSingleCellVignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

# Introduction

This document provides an analysis pipeline for Tabula Muris data using various R packages for single-cell RNA sequencing. Each chunk is annotated with a description of its purpose.

```{r setup, include=FALSE}
# Setup chunk: Load required libraries and set global options
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
```

## Load Required Libraries

The following libraries are loaded for data quality control, visualization, and analysis:

```{r load-libraries, message=FALSE}
# Load necessary libraries for single-cell RNA-seq analysis
library(scater) # Data quality control and visualization
library(BiocParallel) # Parallel computation
library(scran) # Data preprocessing and normalization
library(igraph) # graph handling
library(ggplot2) # visualization
library(SingleCellExperiment) # Core class for single-cell data
library(SingleR) # annotation of cell types
library(celldex) # reference dataset
library(bluster) #
```

## Load Data

For lack of time, we saved the sce in a dedicated object.

```{r, eval=FALSE}
library(SingleCellMultiModal)

mae <- SingleCellMultiModal::scMultiome(dry.run=FALSE)
sce <- mae@ExperimentList$rna
sce
saveRDS(sce, file="inst/extdata/scescmm.rds")
```

Additionally, we load an object with already computed PCA/TSNE/UMAP 
dimentionality reductions.

```{r}
scev2 <- system.file("extdata", "scescmm.rds", package = "scrobinv2")
sce <- readRDS(scev2)
```

## Quality Control

Add quality control metrics to the SingleCellExperiment object:

```{r quality-control}
# Perform quality control and visualize cell metrics
sce <- addPerCellQC(sce)
hist(sce$total, breaks = 50, xlab = "Library Sizes", main = "Distribution of Library Sizes")
hist(sce$detected, xlab="Number of expressed genes", breaks=50,  ylab="Number of cells")

plotHighestExprs(sce, exprs_values = "counts")
```

## Filter Genes with Low Expression
Filter genes based on their average expression levels to exclude non-informative features.

```{r filtering}
wrkrs <- 15
# Calculate average counts for each gene and filter out genes with zero average expression
rowData(sce)$ave.counts <- calculateAverage(sce, exprs_values = "counts", BPPARAM = MulticoreParam(workers = wrkrs))
to.keep <- rowData(sce)$ave.counts > 0
sce <- sce[to.keep, ]
summary(to.keep)  # Summary of genes kept
dim(sce)          # Dimensions of the filtered SingleCellExperiment object

```

## Normalize Counts

Normalize the data using scran's computeSumFactors and log-normalization for downstream analysis.

```{r normcounts}
set.seed(1000)
clusters <- quickCluster(sce, BPPARAM = MulticoreParam(workers = wrkrs))
sce <- computeSumFactors(sce, cluster = clusters, BPPARAM = MulticoreParam(workers = wrkrs))
sce <- logNormCounts(sce)
```

## Variance Modelling and Highly Variable Genes

Model variance to identify highly variable genes for downstream analysis.

```{r hvg}
# Model variance and identify the top highly variable genes
set.seed(1001)
dec <- modelGeneVarByPoisson(sce, BPPARAM = MulticoreParam(workers = wrkrs))
top <- getTopHVGs(dec, prop = 0.1)

```

## Dimensionality Reduction

Perform PCA, t-SNE, and UMAP for dimensionality reduction to visualize the data in lower-dimensional spaces.

For lack of time for GHA, we loaded an object with already computed PCA/TSNE/UMAP.

Here we report the run code.

```{r dimred, eval=FALSE}
# Perform PCA and reduce dimensions using t-SNE and UMAP 
set.seed(1000)
sce <- denoisePCA(sce, subset.row = top, technical = dec, BPPARAM = MulticoreParam(workers = wrkrs))
ncol(reducedDim(sce))  # Number of principal components used
set.seed(1000)
sce <- runTSNE(sce, dimred = "PCA")
set.seed(1000)
sce <- runUMAP(sce, dimred = "PCA")
```

Showing dim reds as TSNE and UMAP.

```{r dimred-plot}
plotReducedDim(sce, dimred = "TSNE", colour_by = "celltype")  # Plot TSNE
plotReducedDim(sce, dimred = "UMAP", colour_by = "celltype")  # Plot UMAP
```


## Graph Construction

Construct a shared nearest neighbor (SNN) graph based on PCA-reduced dimensions.

```{r graph}
# Build a shared nearest neighbor (SNN) graph
graph <- buildSNNGraph(sce, k = 25, use.dimred = 'PCA')  # k=25 neighbors
```

## Cluster Evaluation

Evaluate different clustering methods using robin

```{r}
# Simplify the graph and evaluate clustering methods
graph <- igraph::simplify(graph)  # Simplify the graph by removing duplicate edges
```

# Robin

## Compare all algorithms vs Louvain

We apply the compare procedure to see which is the algorithm that better fits our network.

```{r robin, message=FALSE, warning=FALSE}
graph <- prepGraph(graph, file.format="igraph")
graph <- remove.edge.attribute(graph, "weight")
```


```{r robin2, message=FALSE, warning=FALSE, eval=FALSE}
#Infomap
comp_I <- robinCompare(graph=graph, method1="louvain",
                    method2="infomap")
#Walktrap
comp_W <- robinCompare(graph=graph, method1="louvain",
                       method2="walktrap")
#LabelProp
comp_La <- robinCompare(graph=graph, method1="louvain",
                        method2="labelProp") # use leiden
#Fastgreedy
comp_F <- robinCompare(graph=graph, method1="louvain",
                        method2="fastGreedy")
comp_ll <- robinCompare(graph=graph, method1="louvain",
                        method2="leiden")
```


```{r robin3, message=FALSE, warning=FALSE}
path <- system.file("extdata", package = "scrobinv2")
comp_I <- readRDS(file.path(path, "compI.rds"))
comp_W <- readRDS(file.path(path, "compW.rds"))
comp_La <- readRDS(file.path(path, "compLa.rds"))
comp_F <- readRDS(file.path(path, "compF.rds"))
comp_ll <- readRDS(file.path(path, "compll.rds"))
p <- robin::plotMultiCompare(comp_I,comp_W,comp_F,comp_La, comp_ll, title="Robin Compare") 
p + labs(color = "Model")

```

The lowest curve is the most stable algorithm. 
Leiden and Louvain are the best algorithms in our example. 

## Statistical significance of communities

Due to the fact that Leiden is one of the best algorithm for our network we apply the robust procedure with the Leiden algorithm to see if the communities detected are statistically significant.

```{r, robinrobust, eval=FALSE}
graphRandom <- random(graph=graph)
proc <- robinRobust(graph=graph, graphRandom=graphRandom, method="leiden")
```


```{r robinrobust2}
proc <- readRDS(file.path(path, "proc.rds"))
plot(proc, title="Robin Robust Leiden")

robinFDATest(proc)

robinGPTest(proc)
```

The communities given by Leiden are statistically significant.

## Communities

Here we show a table of the number of cells per each community.

```{r mc}
mc<-membershipCommunities(graph=graph, method="leiden")
table(mc)
```

# Comparing best clustering 

To have a complete overview of the obtained results, we now compare the 
obtained clusters with annotated celltypes by singleR.

To do this kind of analysis, we used as a refernce the `MonacoImmuneData`
dataset provided by the celldex bioconductor package.

```{r annotation, message=FALSE}
monacosce <- MonacoImmuneData()

sr <- SingleR(test=sce, ref=monacosce, labels=monacosce$label.fine)

sce$singleR <- sr$labels
```

We now store the communities into the sce object and use them in combination with the reference to annotate them.

```{r communities, message=FALSE}
set.seed(999)

sce$leiden <- as.factor(mc)

src <- SingleR(test=sce, ref=monacosce, labels=monacosce$label.fine, clusters=sce$leiden)
names(src$labels) <- names(table(sce$leiden))

sce$singleRLeiden <- NA
for(i in names(src$labels))
{
    sce$singleRLeiden[which(sce$leiden == i)] <- src$labels[i]
}
```

Here we compare the results obtained between the original cell types and the 
annotated results using the transformed VI metric.

```{r vinorm}
## Comparison between ground truth and SingleR labels alone
igraph::compare(sce$celltype, sce$singleR, method="vi")/log2(dim(sce)[2]) # vi=1.405753 
## Comparison between ground truth and Clusters computed with Leiden
igraph::compare(sce$celltype, sce$leiden, method="vi")/log2(dim(sce)[2]) # vi=1.067032
## Comparison between ground truth and SingleR with labels + clusters
igraph::compare(sce$celltype, sce$singleRLeiden, method="vi")/log2(dim(sce)[2]) # 0.9767742
```

Finally, to better investigate the results, we plot a heatmap where we have the 
original cell types on the y axis and the different annotations on the x axis.

```{r plot-clusters}
source(system.file("scripts", "functions.R", package = "scrobinv2"))
df <- as.data.frame(colData(sce))
ClusterHeatmap(df, "celltype", "leiden")
ClusterHeatmap(df, "celltype", "singleR")
ClusterHeatmap(df, "celltype", "singleRclust")
```

# Session Info

```{r, tidy=TRUE}
sessionInfo()
```


