---
title: "1. Seurat Robin Single Cell Vignette"
author: "Valeria Policastro"
date: "`r Sys.Date()`"
output: 
    html_document:
        code_folding: hide
        toc: true
        toc_float: true
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{RobinSingleCellVignette}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include=FALSE}
### SET GENERAL OPTIONS FOR ALL CODE CHUNKS!
knitr::opts_chunk$set(echo = TRUE, message=FALSE, cache=TRUE, warning=FALSE)
get_data_path <- function(fname) {
    local_path <- file.path("..", "inst", "extdata", fname)
    if (file.exists(local_path)) {
        return(local_path)
    }
    local_path <- file.path("..", "inst", "scripts", fname)
    if (file.exists(local_path)) {
        return(local_path)
    }
    pkg_path <- system.file("extdata", fname, package = "scrobinv2")
    if (nzchar(pkg_path)) {
        return(pkg_path)
    }
    pkg_path <- system.file("scripts", fname, package = "scrobinv2")
    if (nzchar(pkg_path)) {
        return(pkg_path)
    }
    stop("File not found: ", fname)
}
```

# Single cell RNAseq Analysis

## Package loading 

```{r, message=FALSE}

library(dplyr)
library(Seurat)
library(patchwork)
library(mclust)
library(robin)
```

## Reading data 

```{r}
# datav2 <- system.file("extdata", "filtered_tabulamuris_mtx.rds", package="scrobinv2")
datav2 <- get_data_path("filtered_tabulamuris_mtx.rds")
tabData <- readRDS(file=datav2)

SingleCellData <- CreateSeuratObject(counts = tabData, project = "singleCell", min.cells = round(dim(tabData)[2]*5/100), min.features = 0)
SingleCellData 
```

## Data normalization

```{r}

SingleCellData <- NormalizeData(SingleCellData, normalization.method = "LogNormalize", scale.factor = 10000)

SingleCellData[["RNA"]]@layers$counts[1:5,1:5]


```

## Feature selection 

```{r}

SingleCellData <- FindVariableFeatures(object = SingleCellData, selection.method="vst", nfeatures = 2000)


```

## Scale the data 

```{r}
SingleCellData <- ScaleData(SingleCellData, do.center = TRUE, do.scale = TRUE)
SingleCellData[["RNA"]]@layers$scale.data[1:5,1:5]
```


## Linear dimension reduction

```{r}

SingleCellData <- RunPCA(SingleCellData, features = VariableFeatures(object = SingleCellData))
```


## Graph 

```{r}

SingleCellData <- FindNeighbors(SingleCellData, dims = 1:50) 

SingleCellData@graphs$RNA_snn[1:5,1:5] # Adjacency matrix
```


# How to use Robin

## Reading Graph

```{r}


AdjSNN <- SingleCellData@graphs$RNA_snn
graphSingleCell <- graph_from_adjacency_matrix(AdjSNN,mode="directed",weighted = TRUE,add.colnames = "NA",diag=FALSE)
edge <- as_edgelist(graphSingleCell)
graph <- igraph::graph_from_edgelist(edge, directed=FALSE)
graph <- igraph::simplify(graph)
graph

```



## Compare all algorithms vs Louvain

We apply the compare procedure to see which is the algorithm that better fits our network.

```{r}
#Infomap
 comp_I <- robinCompare(graph=graph, method1="louvain",
                        method2="infomap")
 plot1 <- plot(comp_I)


#Walktrap
comp_W <- robinCompare(graph=graph, method1="louvain",
                       method2="walktrap")

plot2 <- plot(comp_W)


#LabelProp
comp_La <- robinCompare(graph=graph, method1="louvain",
                        method2="labelProp")


plot3 <- plot(comp_La)

#Fastgreedy
comp_F <- robinCompare(graph=graph, method1="louvain",
                        method2="fastGreedy")

plot4 <- plot(comp_F)


PlotComparisonAllVsInfomap <- gridExtra::grid.arrange(plot1,plot2,plot3,plot4, ncol=2)

```


The lowest curve is the most stable algorithm. Louvain and Walktrap are the best algorithms in our example. 



## Statistical significance of communities

Due to the fact that Louvain is one of the best algorithm for our network we apply the robust procedure with the Louvain algorithm to see if the communities detected are statistically significant.

```{r}
graphRandom <- random(graph=graph)
proc <- robinRobust(graph=graph, graphRandom=graphRandom, method="louvain")

plot(proc)

robinFDATest(proc)

robinGPTest(proc)
```

The communities given by Louvain are statistically significant.

# Differential Analysis  (Marker Gene Identification)

Once we found the best clustering algorithm, we can use the found communities
for identifying marker genes, to subsequently find cell types.

```{r, message=FALSE}
mc <- membershipCommunities(graph=graph, method="louvain")
SingleCellData$seurat_clusters = mc
Idents(SingleCellData) <- "seurat_clusters"
markers <- FindAllMarkers(SingleCellData, only.pos = TRUE, logfc.threshold = 0.25)

markers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 1) %>%
    slice_head(n = 10) %>%
    ungroup() -> top10
DoHeatmap(SingleCellData, features = top10$gene) + NoLegend()
```





# SessionInfo
```{r, tidy=TRUE}
sessionInfo()

```

